---
title: "DailyTest"
author: "Rainier Pederson"
date: "2026-09-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(UsingR)
```

## 1. What categories are there?

There are ways to check how many unique values there are in the Groups section of the data. None of them are as slow as trying to re-learn R from first principles, but none of them get me to re-learn R from first principles.

To start, the table:

```{r}
jons_guest_list <- read.csv("C:/Users/Grad School/Desktop/RCode/Data/daily_show_guests.csv", 
                            header = TRUE)
head(jons_guest_list, 10)
```

With that code having run, we can see the types of data in the file:

```{r}
colnames(jons_guest_list)
```

From there, I'm interested in the Groups:

```{r}
jons_groups <- jons_guest_list$Group
jons_groups <- as.data.frame(jons_groups)
jons_groups <- distinct(jons_groups)
jons_groups
```

Now we know what the different groups are. Yay!


## 2. Plot change over time

I don't care about any data beyond the year and group:

```{r}
jons_guest_list <- subset(jons_guest_list, select = c("YEAR", "Group"))
head(jons_guest_list, 10)
```

Rotating that so we have a new in which each column is a year, each row is a profession (Group), and that intersection is how many of that profession Jon had on that year: 

```{r}
jons_guest_list <- count(jons_guest_list, Group, YEAR)
#jons_guest_list
# jons_guest_list <- pivot_wider(jons_guest_list, names_from = YEAR, 
                               #values_from = n, values_fill = 0)

head(jons_guest_list, 10)
```

So with all that done, I should be able to finally build this graph:

```{r}
ggplot(jons_guest_list, aes(fill=Group, y=n, x=YEAR)) + 
    geom_bar(position="stack", stat="identity")
```

I skipped a data cleaning step earlier, so I'm making up for it now. Namely: 
1. "N/A" indicates no guest should be discounted;
2. "media" and "Media" should be the same under "Media";
3. "Academic" and "Science" can be merged;
4. "Advocacy", "Business", "Clergy", "Consultant", "Military", and "Political Aide" can all go under Misc.
5. The chart should be normalized because if I'm comparing how popular politicians v comedians are, it isn't helpful to say they both went down in popularity by about half when Jon left the show halfway through the year. 
6. The y-axis should be called something sensible.

For the N/A:

```{r}
jons_guest_list <- filter(jons_guest_list, Group != "N/A")
ggplot(jons_guest_list, aes(fill=Group, y=n, x=YEAR)) + 
    geom_bar(position="stack", stat="identity")
```


For the rest, it's easier to just start with clean data

```{r}
jons_guest_list_2 <- read.csv("C:/Users/Grad School/Desktop/RCode/Data/daily_show_guests.csv", 
                              header = TRUE)
jons_guest_list_2 <- filter(jons_guest_list_2, Group != "N/A")
jons_guest_list_2$Group[jons_guest_list_2$Group == "media"] <- 
                                                   "Media"
jons_guest_list_2$Group[jons_guest_list_2$Group == "Academic"] <-
                                                   "Academic/Science"
jons_guest_list_2$Group[jons_guest_list_2$Group == "Science"] <- 
                                                   "Academic/Science"
jons_guest_list_2$Group[jons_guest_list_2$Group == "Advocacy"] <- 
                                                   "Misc"
jons_guest_list_2$Group[jons_guest_list_2$Group == "Business"] <- 
                                                   "Misc"
jons_guest_list_2$Group[jons_guest_list_2$Group == "Clergy"] <- 
                                                   "Misc"
jons_guest_list_2$Group[jons_guest_list_2$Group == "Consultant"] <- 
                                                   "Misc"
jons_guest_list_2$Group[jons_guest_list_2$Group == "Military"] <- 
                                                   "Misc"
jons_guest_list_2$Group[jons_guest_list_2$Group == "Political Aide"] <- 
                                                   "Misc"
jons_guest_list_2 <- subset(jons_guest_list_2, select = c("YEAR", "Group"))
jons_guest_list_2 <- count(jons_guest_list_2, Group, YEAR)

ggplot(jons_guest_list_2, aes(fill=Group, y=n, x=YEAR)) +
    geom_bar(position="fill", stat="identity")
```

So that's basically what I want. To make this a stacked line chart, and then add some custom coloring, and then actually deciding to not normalize the data:


```{r}
custom_colors = c("#d14d4b", "#ecb87c", "#a78c11", "#cfee51", "#629938", "#b5ffe1", "#2d7f91", "#9a92da", "#4a1554")

ggplot(jons_guest_list_2, aes(fill=Group, y=n, x=YEAR)) +
    geom_area() + 
    labs(
          title = "The Types of Guests on Jon Stwert's Daily Show Over Time",
          subtitle = "Data procured from FiveThirtyEight",
          caption = "Graph by Rainier Pederson. Data from FiveThirtyEight,\n 
          found at https://github.com/fivethirtyeight/data/tree/master/daily-show-guests.\n 
          Used under CC by 4.0.",
          tag = waiver(),
          dictionary = waiver(),
          alt = waiver(),
          alt_insight = waiver(),
          x = "Year",
          y = "Number of Guests"
        ) +
  scale_fill_manual(values = custom_colors)

```

There are weird jumps that correspond to a lack of data where data should have been set to 0. This wasn't caught in the bar graph stage because there it wouldn't have mattered. To clean that:

```{r}
head(jons_guest_list_2, 10)
jons_guest_list_3 <- jons_guest_list_2
jons_guest_list_2 <- pivot_wider(jons_guest_list_2, names_from = YEAR, 
                                 values_from = n, values_fill = 0)
jons_guest_list_2 <- pivot_longer(jons_guest_list_2, !Group, names_to = "YEAR")
# print(as_tibble(jons_guest_list_2), n = 200)
# Yes, the 0s are 

# jons_guest_list_2 <- as.data.frame(jons_guest_list_2)
jons_guest_list_2$YEAR <- as.integer(jons_guest_list_2$YEAR)

ggplot(jons_guest_list_2, aes(fill=Group, y=value, x=YEAR)) +
    geom_area() + 
    labs(
          title = "The Types of Guests on Jon Stwert's Daily Show Over Time",
          subtitle = "Data procured from FiveThirtyEight",
          caption = "Graph by Rainier Pederson. Data from FiveThirtyEight,\n found at https://github.com/fivethirtyeight/data/tree/master/daily-show-guests.\n Used under CC by 4.0.",
          tag = waiver(),
          dictionary = waiver(),
          alt = waiver(),
          alt_insight = waiver(),
          x = "Year",
          y = "Number of Guests"
        ) +
  scale_fill_manual(values = custom_colors)
```

That took a minute (week) to work. As it turns out, when jons_guest_list_2 was pivoting and un-pivoting, the YEAR went from int to char and that caused a whole bunch of problems. 

Anyways, my last thing is to run the data but without including 2015 because visually, the sudden decrease is throwing me.

```{r echo=FALSE}
ggplot(filter(jons_guest_list_2, YEAR != 2015), aes(fill=Group, y=value, x=YEAR)) +
    geom_area() + 
    labs(
          title = "The Types of Guests on Jon Stwert's Daily Show Over Time \n
          From 1999-2014",
          subtitle = "Data procured from FiveThirtyEight",
          caption = "Graph by Rainier Pederson. Data from FiveThirtyEight,\n found at https://github.com/fivethirtyeight/data/tree/master/daily-show-guests.\n Used under CC by 4.0.",
          tag = waiver(),
          dictionary = waiver(),
          alt = waiver(),
          alt_insight = waiver(),
          x = "Year",
          y = "Number of Guests"
        ) +
  scale_fill_manual(values = custom_colors)
```

And finally, exporting just the image to show my dad:

```{r}
ggsave("DailyGuests1999-2014.png")
```

And with all that, I think this has been a successful first project in getting back into R! 